---
title: "Appendix - Visualizing thefts and robberies on transportation in Mexico City with R"
author: "Rebeca de Buen Kalman"
date: "5/15/2020"
output: html_document
---

```{r, warning = FALSE, message = FALSE}
library(readr) # reading in data
library(dplyr) # data manipulation
library(forcats) # recoding variables
library(ggplot2) # visualization
library(lubridate) # date-time object manipulation
library(stringr) # strings and regular expressions
library(zoo) # moving average functions
library(sf) # working with spatial objects
```

```{r}
load(file = "../processed_data/transport_crimes.Rdata")
```
In this Appendix I show the code for the plots in the post *A quick glance at reported robberies and thefts on Mexico City's transportation system*. In the previous Apendix I laied out the process followed for preparing the data.

## Research questions
Before jumping into the data, I brainstormed a set of questions to guide this analysis:

- What kinds of crimes are reported on Mexico City's transportation system (both private cars and public transportation)?
- How do the crimes reported vary by mode?
- Have reported crimes increased  or decreased over time? Does this vary by mode?
- What are the  months, days of the week and hours when crimes are more frequently reported on different modes of transportation?
- For crimes committed on the Metro and Metrubus what are the most common stations committed?
- For other modes, what neighborhoods are the most dangerous?
- For most problematic neighborhoods, what are the days and times when more crimes are reported?

## Analysis and visualization

 *What kinds of crimes are committed on Mexico City's transportation system (both private cars and public transportation)?How do the crimes reported vary by mode?*

The crimes included in the data set for modes of transportation are violent robberies and non-violent thefts. I explore how these are distributed across modes for the entire time period (2016-2019).

```{r, warning = FALSE, message = FALSE}
# I group by mode and count so that my categorical variable is ordered by the frequency of crimes
mode_levels <- transport_crimes %>% group_by(transp_mode) %>% count() %>% arrange(n) %>% pull(transp_mode)

# I make a simple frequency plot to vizualize how many crimes are reported for each mode. I separate these by violent and non-violent crime.
transport_crimes %>% select(transp_mode, violence) %>% 
                      #order frequency plot modes from highest to lowest counts
                      mutate(transp_mode = factor(transp_mode, levels=rev(mode_levels))) %>% 
                      ggplot(aes(x=transp_mode, fill = violence)) + 
                      geom_bar() + 
                      # fix labels and legend
                      labs(
                      title = "Number of reported robberies and thefts by mode of transport",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      xlab("Mode") +
                      ylab("Number of reports") +
                      scale_fill_discrete(name = "Violent?") +
                      theme_minimal() +
                      theme(legend.position= c(0.2,-0.12), legend.direction="horizontal") +
                      theme(axis.text.x = element_text(angle = 30)) +
                      theme(axis.title.x = element_blank())
                    
  
``` 
Now by year:

```{r, warning = FALSE, message = FALSE}
#totals <- transport_crimes %>% group_by(year) %>% count()

year <- c(2016, 2017, 2018, 2019) 
x_coords <- c(5, 5, 5, 5)
y_coords <- c(3500, 3500, 3500, 3500)
text <- c("Total: 4,316", 
          "Total: 6,838", 
          "Total: 12,440",
          "Total: 12,292")

year_totals <- data.frame(year, x_coords, y_coords, text)

transport_crimes %>% select(transp_mode, violence, year) %>% 
                      #order frequency plot modes from highest to lowest counts
                      mutate(transp_mode = factor(transp_mode, levels=rev(mode_levels))) %>% 
                      ggplot(aes(x=transp_mode, fill = violence)) + 
                      geom_bar() + 
                      geom_text(data = year_totals, 
                                mapping = aes(x = x_coords, y = y_coords, label = text), 
                                inherit.aes = FALSE, size = 3) +
                      facet_wrap(~year) +
                      # fix labels and legend
                      labs(
                      title = "Number of reported robberies by mode of transport",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      xlab("Mode") +
                      ylab("Number of reports") +
                      scale_fill_discrete(name = "Violent?") +
                      theme_minimal() +
                      theme(legend.position= c(0.2,-0.13), legend.direction="horizontal") +
                      theme(axis.text.x = element_text(angle = 30)) +
                      theme(axis.title.x = element_blank())
                                
  
```

* Have reported crimes increased  or decreased over time? Does this vary by mode?*

```{r, warning = FALSE, message = FALSE}
transport_crimes %>%  
  select(transp_mode, violence, day) %>% 
  group_by(day, violence, transp_mode) %>% count() %>% ungroup() %>%
  group_by(transp_mode, violence) %>%
  arrange(day) %>% # Make sure days are in order for rolling mean
  mutate(avg_n_7 = rollmean(n, 14, fill = c(NA, "extend", NA)))  %>% # 14 day rolling mean
                    ggplot(aes(x = day, y = avg_n_7, color = violence, group = violence)) + geom_line() + 
                    facet_grid(transp_mode ~ .) +
                    scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
                    ylab("Number of reports") + xlab("Time") +
                    labs(
                    title = "Number of daily reported robberies and thefts over time by mode of transport",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      scale_color_discrete(name = "Violent?") +
                    theme_minimal() +
                    theme(strip.text = element_text(size = 7)) +
                    theme(axis.title.x = element_blank()) + 
                    theme(legend.position= c(0.2,-0.21), legend.direction="horizontal") +
                    theme(axis.text.x = element_text(angle = 90)) 

```
 
```{r, warning = FALSE, message = FALSE}
 transport_crimes %>% select(violence, year, month, transp_mode) %>% 
                    group_by(year, month, transp_mode) %>%
                    count(month) %>%
                    ggplot(aes(x= month, y = n, group = year, color = year))  + geom_line() +
                    facet_wrap(~transp_mode)  +
                    labs(title = "Number of reported robberies and thefts by month and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      xlab("Month") +
                      ylab("Number of reports") + 
                    scale_x_discrete(labels=c("1" = "January",
                                              "2" = "February", 
                                              "3" = "March",
                                              "4" = "April", 
                                              "5" = "May", 
                                              "6" = "June", 
                                              "7" = "July",
                                              "8" = "August",  
                                              "9" = "September" , 
                                              "10" = "October", 
                                              "11" = "November", 
                                              "12" = "December"))  + 
                    theme_minimal() +
                    theme(axis.text.x = element_text(angle = 90)) + 
                    theme(legend.position= c(0.2,-0.215), legend.direction="horizontal") +
                    theme(axis.title.x = element_blank())
```

*What are the  months, days of the week and hours when crimes are more frequently committed on different modes of transportation?*

```{r, warning = FALSE, message = FALSE}
transport_crimes %>% select(violence, month, transp_mode, year) %>% 
                    group_by(violence, month, transp_mode, year) %>%
                    ggplot(aes(x= month, group = violence, fill = violence))  + 
                    geom_bar() +
                    facet_grid(transp_mode ~ year) +
                    labs(title = "Number of reported robberies and thefts by month and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      ylab("Number of reports") + 
                    scale_x_discrete(labels=c("1" = "January", 
                                              "2" = "February", 
                                              "3" = "March",
                                              "4" = "April", 
                                              "5" = "May", 
                                              "6" = "June", 
                                              "7" = "July",
                                              "8" = "August", 
                                              "9" = "September" , 
                                              "10" = "October", "11" =
                                                "November", 
                                              "12" = "December"))  + 
                    scale_fill_discrete(name = "Violent?") + theme(axis.text=element_text(size=7)) +
                    theme_minimal() + 
                    theme(axis.text.x = element_text(angle = 90)) + 
                    theme(legend.position= c(0.2,-0.20), legend.direction="horizontal") +
                    theme(axis.title.x = element_blank())
```


```{r, warning = FALSE, message = FALSE}
transport_crimes %>% select(transp_mode, violence, weekday, year) %>%
                    group_by(violence, weekday, transp_mode) %>%
                    ggplot(aes(x= weekday, group = violence, fill = violence))  +
                    geom_bar() + 
                    facet_grid(year~transp_mode)  +
                      labs(title = "Number of reported robberies and thefts by weekday and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      ylab("Number of reports") +
                    scale_fill_discrete(name = "Violent?") +
                    scale_x_discrete(labels=c("1" = "Sunday",
                                              "2" = "Monday",
                                              "3" = "Tuesday",
                                              "4" = "Wednesdy", 
                                              "5" = "Thursday", 
                                              "6" = "Friday", 
                                              "7" = "Saturday"))  +
                    scale_fill_discrete(name = "Violent?") +
                    theme_minimal() +
                    theme(axis.title.x = element_blank()) +
                    theme(strip.text = element_text(size = 7)) + 
                    theme(axis.text.x = element_text(angle = 90)) + 
                    theme(legend.position= c(0.15,-0.23), legend.direction="horizontal") 
```

```{r, warning = FALSE, message = FALSE}
 transport_crimes %>%  select(transp_mode, datetime, violence) %>% 
                    mutate(hour = hour(datetime), year = year(datetime)) %>% 
                    ggplot(aes(x = hour, fill = violence)) + geom_bar() +
                    facet_grid(year~transp_mode) +
                      labs(title = "Number of reported robberies and thefts by hour of day and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      ylab("Number of reports") + 
                      scale_x_continuous(breaks = c(0,5,10,15,20), labels= c(`0` = "00:00", `5` = "5:00", 
                                                                             `10` = "10:00", `15`= "15:00", 
                                                                             `20` = "20:00")) + 
                    scale_fill_discrete(name = "Violent?") +
                    theme_minimal() +
                    theme(axis.text.x = element_text(angle = 90)) +
                    theme(axis.title.x = element_blank()) + 
                    theme(axis.text=element_text(size=7)) + 
                    theme(legend.position= c(0.15,-0.12), legend.direction="horizontal") +
                    theme(strip.text.x = element_text(size = rel(0.6))) 
```


*For crimes committed on the Metro what are the most common stations where they are committed?*

To explore questions related to locations, I will use the additional data sets that include shapefiles for Mexico City Neighborhoods, Transit (Metro and Metrobus) Lines, and Transit Stations. 

```{r, warning = FALSE, message = FALSE}
# Load neighborhood ("colonias") shapefiles
colonias <- st_read("../data/coloniascdmx.shp", quiet = TRUE) %>% 
  select(name = nombre, geometry) %>% 
  st_transform(4485)

# glance at the data
head(colonias)
```

```{r, warning = FALSE, message = FALSE}
# make lat and lon into gemetrical object sf can work with
transport_crimes_geolocated <- transport_crimes %>% 
  filter(!is.na(lat)) %>% 
  st_as_sf(coords = c("lon", "lat"), crs=4326) %>% 
  st_transform(4485)
```

```{r, warning = FALSE, message = FALSE}
metro_lines <- st_read("../data/lineas-de-metro/lineas-de-metro.shp", quiet = TRUE) %>% st_transform(4485) %>% 
                    mutate(line = fct_recode(name, "Line 1" = "Línea 1", 
                                               "Line 2" = "Línea 2",
                                               "Line 3" = "Línea 3",
                                               "Line 4" = "Línea 4",
                                               "Line 5" = "Línea 5",
                                               "Line 6" = "Línea 6",
                                               "Line 7" = "Línea 7",
                                               "Line 8" = "Línea 8",
                                               "Line 9" = "Línea 9",
                                               "Line A" = "Línea A",
                                               "Line B" = "Línea B",
                                               "Line 12" = "Línea 12")) %>% 
                                              select(-name)
                           
metro_stations <- st_read("../data/estaciones-metro/estaciones-metro.shp", quiet = TRUE) %>%
  st_transform(4485) %>% 
  st_buffer(dist = units::as_units(250, "m"))
```

Plot the points where crimes are reported over the metro lines
```{r, warning = FALSE, message = FALSE}

metro_crimes <- transport_crimes_geolocated %>%
  filter(transp_mode == "Metro")
cols <- c("Line 1" = "hotpink1", 
          "Line 2" = "blue", 
          "Line 3" = "olivedrab", 
          "Line 4" = "cadetblue1", 
          "Line 5" = "yellow",
          "Line 6" = "red",
          "Line 7" = "orange", 
          "Line 8" = "darkgreen", 
          "Line 9" = "brown", 
          "Line A" = "purple", 
          "Line B" = "grey69", 
          "Line 12" = "darkgoldenrod")

violence.labs <- c(Yes = "Violent", No = "Non-violent")

ggplot(colonias) + 
  geom_sf(size = 0.1) + 
  geom_sf(data = metro_lines, aes(color = line), size = 1) + 
  scale_colour_manual(values = cols) + 
  geom_sf(data=metro_crimes,  color = "black", size = 0.4, alpha = 0.5) +
    facet_wrap(~violence, labeller = labeller(violence = violence.labs)) +
  theme_void() + theme(legend.title = element_blank()) +
  theme(legend.position= "bottom", 
        legend.direction="horizontal")
```


```{r}
colonias_crimes <- colonias %>% st_join(metro_crimes) %>% filter()

colonias_counts <- colonias_crimes %>% 
  filter(!is.na(name)) %>% 
  st_drop_geometry() %>% 
  group_by(name) %>% 
  summarize(n = n()) %>% 
  full_join(colonias %>% 
              filter(!is.na(name)) %>%
              group_by(name) %>% 
              slice(1L) %>% 
              select(name, geometry)) %>%
  mutate(n = ifelse(is.na(n), 0, n)) %>%
  st_as_sf()

colonias_counts %>% ggplot(aes(fill = n)) + 
  geom_sf(size = 0.1) + 
  scale_fill_gradient(low = "white", high = "red", 
                      na.value = NA, trans = "log", 
                      breaks = c(1, 7, 50, 400)) + theme_void() +
    geom_sf(data = metro_lines, aes(color = line), 
            size = 0.5, inherit.aes = FALSE) +
  scale_colour_manual(values = cols) +
     theme(legend.title = element_blank()) +
  theme(legend.position= "bottom", legend.direction="horizontal")
```

```{r, warning = FALSE, message = FALSE}
station_crimes <- metro_crimes %>% st_join(metro_stations)
station_counts <- station_crimes %>% 
  st_drop_geometry() %>% 
  group_by(violence, stop_name) %>% 
  summarize(n = n()) %>% arrange(desc(n, by_group = TRUE)) %>%
  mutate(stop_name = str_remove_all(stop_name, "[_1234567890]")) %>% 
  ungroup()

top_violent_stations <- station_counts %>% select(stop_name, violence, n) %>% 
  filter(!is.na(stop_name), violence == "Yes") %>% 
  group_by(stop_name) %>% slice(1)  %>% 
  arrange(desc(n, by_group = TRUE)) %>% 
  ungroup() %>% 
  select(-violence) %>% 
  top_n(10) 


top_nonviolent_stations <- station_counts %>% 
  select(stop_name, violence, n) %>% 
  filter(!is.na(stop_name), violence == "No") %>% 
  group_by(stop_name) %>% 
  slice(1)  %>% 
  arrange(desc(n, by_group = TRUE)) %>% 
  select(-violence) %>% 
  top_n(10)
  

```



```{r}

# Function to count number of reports by neighborhood and prepare for map 
mode_crimes_counts <- function(mode){
  #filter out the reports of the chosen mode
mode_crimes <- transport_crimes_geolocated %>% filter(transp_mode == mode)
# spatial join with the neihborhoods data
colonias_crimes <- colonias %>% st_join(mode_crimes)
# count reports by neighborhood
colonias_counts <- colonias_crimes %>% 
  filter(!is.na(name)) %>% 
  st_drop_geometry() %>% 
  group_by(name) %>% 
  summarize(n = n()) %>% 
  full_join(colonias %>% 
              filter(!is.na(name)) %>%
              group_by(name) %>% 
              slice(1L) %>% 
              select(name, geometry)) %>%
  mutate(n = ifelse(is.na(n), 0, n)) %>%
  st_as_sf()
return(colonias_counts)
}
```

```{r}
mode_crimes_counts("Car")
```

```{r, warning = FALSE, message = FALSE, echo=FALSE}

# Function to count top 10 neighborhoods for a given mode
top_10_neighborhoods <- function(mode){
  top_10 <- mode_crimes_counts(mode) %>% 
    st_drop_geometry() %>% 
    arrange(desc(n, by_group = TRUE)) %>% 
    top_n(10)
  return(top_10)
}
```

```{r}
top_10_neighborhoods("Metro")

```



```{r}

# Function to map neighborhood reports for a given mode
mode_crimes_map <- function(mode){
  colonias_counts <- mode_crimes_counts(mode)
colonias_counts %>% ggplot(aes(fill = n)) + 
  geom_sf(size = 0.1) + 
  scale_fill_gradient(low = "white", high = "red", na.value = NA, 
                      trans = "log", breaks = c(1, 30, 100))  + 
  theme_void() + 
   theme(legend.title = element_blank()) +
  labs(title = paste0("Number of reported robberies and thefts by ", mode, " users"),
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México")
}
```


```{r}
top_10_neighborhoods("Pesero")

```

```{r}
#compute a different way (colonias or neighborhoods have slightly different names and divisions in shapefiles than admin report)
transport_crimes %>% filter(transp_mode == "Car") %>% 
  group_by( municipality) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10)
```

```{r}
mode_crimes_map("Car")
```

```{r}
mode_crimes_map("Pesero")
```