---
title: "A quick glance at reported robberies and thefts in Mexico City’s transportation system"
author: "Rebeca de Buen Kalman"
date: "5/8/2020"
output: 
  html_document:
    theme: journal
---

```{r, warning = FALSE, message = FALSE, echo = FALSE}
library(readr) # reading in data
library(dplyr) # data manipulation
library(forcats) # recoding variables
library(ggplot2) # visualization
library(lubridate) # date-time object manipulation
library(stringr) # strings and regular expressions
library(zoo) # moving average functions
library(sf) # working with spatial objects
```

```{r, warning = FALSE, message = FALSE, echo = FALSE}
load(file = "../processed_data/transport_crimes.Rdata")
```


When the Mexico City Government published records of street level crime in 2019, I became curious about what these data might show about crime on different modes of transportation. Last week, I finally dived into this data set (because anything to get my mind off COVID-19 please) and here is what I learned when I analyzed four full years of crime reports, from 2016-2019.

#### The crime reports on modes of transportation are about violent robberies and non-violent thefts. Mass transit users report frequent thefts but most reports of crime on Metro and Metrobus are non-violent. Drivers report the most violent crime followed by peseros and Taxis.

The mode of transportation where people report the most violent crime is consistently the car. While mass transit users (Metro and Metrobus) also report large amounts of thefts, most of the reports from transit users are reported as non-violent. Pesero and Taxi users report less overall crimes but most of the reports of crime on these modes include violence.

The total amount of reports has consistently grown since 2016. However, it is hard to say whether this increase in reports is because there is more crime or because people have become more likely to file a report. The large proportion of reports of robberies commited against drivers and passengers of cars may indicate that car users are more likely to report crimes committed against them. Similarly, users of other modes like the Pesero and other forms of public transportation may be less likely to report robberies against them to the police. Finally, other types of crimes, like sexual assault, are committed on public transportation but are likely classified in a way that does not make it clear that it occured on a mode of transportation. Therefore, this analysis captures only crimes that are explicitely reported as crimes on transportation.


```{r, warning = FALSE, message = FALSE, echo = FALSE}
# I group by mode and count so that my categorical variable is ordered by the frequency of crimes
mode_levels <- transport_crimes %>% group_by(transp_mode) %>% count() %>% arrange(n) %>% pull(transp_mode)

year <- c(2016, 2017, 2018, 2019) 
x_coords <- c(5, 5, 5, 5)
y_coords <- c(3500, 3500, 3500, 3500)
text <- c("Total: 4,316", 
          "Total: 6,838", 
          "Total: 12,440",
          "Total: 12,292")

year_totals <- data.frame(year, x_coords, y_coords, text)

transport_crimes %>% select(transp_mode, violence, year) %>% 
                      #order frequency plot modes from highest to lowest counts
                      mutate(transp_mode = factor(transp_mode, levels=rev(mode_levels))) %>% 
                      ggplot(aes(x=transp_mode, fill = violence)) + 
                      geom_bar() + 
                      geom_text(data = year_totals, 
                                mapping = aes(x = x_coords, y = y_coords, label = text), 
                                inherit.aes = FALSE, size = 3) +
                      facet_wrap(~year) +
                      # fix labels and legend
                      labs(
                      title = "Number of reported robberies and thefts by mode",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      xlab("Mode") +
                      ylab("Number of reports") +
                      scale_fill_discrete(name = "Violent?") +
                      theme_minimal() +
                      theme(legend.position= c(0.2,-0.13), legend.direction="horizontal") +
                      theme(axis.text.x = element_text(angle = 30)) +
                      theme(axis.title.x = element_blank())
                                
  
```


#### Reports of crime on transportation generally increased in 2018 but have mostly decreased since early 2019.

There is large variation in the number of reports over time. If taken at face value, the data indicates that there was a large increase in robberies through 2018 and into 2019. However, reported crimes decreased to a close to all-time low point by the end of 2019.


```{r, warning = FALSE, message = FALSE, echo = FALSE}
transport_crimes %>%  
  select(transp_mode, violence, day) %>% 
  group_by(day, violence, transp_mode) %>% count() %>% ungroup() %>%
  group_by(transp_mode, violence) %>%
  arrange(day) %>% # Make sure days are in order for rolling mean
  mutate(avg_n_7 = rollmean(n, 14, fill = c(NA, "extend", NA)))  %>% # 14 day rolling mean
                    ggplot(aes(x = day, y = avg_n_7, color = violence, group = violence)) + geom_line() + 
                    facet_grid(transp_mode ~ .) +
                    scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
                    theme_minimal() +
                    ylab("Number of reports") + xlab("Time") +
                    theme_minimal() +
                    theme(axis.text.x = element_text(angle = 90)) +
                      labs(
                      title = "Number of daily reported robberies and thefts over time by mode",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      theme(legend.position= c(0.2,-0.21), legend.direction="horizontal") +
                      scale_color_discrete(name = "Violent?") +
                      theme(strip.text = element_text(size = 7)) +
                       theme(axis.title.x = element_blank())

```

#### There is no clear pattern that indicates that any particular month of the year is more dangerous for any particular mode of transportation.

The reports of robberies on all modes of transportation vary randomly across months for all of the years analyzed. There does not appear to be any month of the year where robberies are consistently higher than other months.

```{r, warning = FALSE, message = FALSE, echo = FALSE}
transport_crimes %>% select(violence, month, transp_mode, year) %>% 
                    group_by(violence, month, transp_mode, year) %>%
                    ggplot(aes(x= month, group = violence, fill = violence))  + 
                    geom_bar() +
                    facet_grid(transp_mode ~ year) +
                    labs(title = "Number of reported robberies by month and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      ylab("Number of reports") + 
                    scale_x_discrete(labels=c("1" = "January", "2" = "February", "3" = "March",
                                              "4" = "April", "5" = "May", "6" = "June", "7" = "July",
                                              "8" = "August",  "9" = "September" , "10" = "October", "11" =
                                                "November", "12" = "December"))  + 
                    theme_minimal() + 
                    theme(axis.text.x = element_text(angle = 90)) + 
                    theme(legend.position= c(0.15,-0.20), legend.direction="horizontal") +
                    scale_fill_discrete(name = "Violent?") + theme(axis.text=element_text(size=7)) +
                    theme(axis.title.x = element_blank()) +
                      theme(strip.text = element_text(size = 7))
```

#### Robberies are consistently lowest on Sundays followed by Saturdays

Accross all modes and for all of the years analyzed, robberies are less frequently reported on weekends. This is not surprising since people are also less likely to travel on weekends, especially Sundays which is the day with the lowest amount of reports.


```{r, warning = FALSE, message = FALSE, echo = FALSE}
transport_crimes %>% select(transp_mode, violence, weekday, year) %>%
                    group_by(violence, weekday, transp_mode) %>%
                    ggplot(aes(x= weekday, group = violence, fill = violence))  +
                    geom_bar() + 
                    facet_grid(year~transp_mode) +
                    theme_minimal() +
                    scale_x_discrete(labels=c("1" = "Sunday", "2" = "Monday", "3" = "Tuesday",
                                              "4" = "Wednesdy", "5" = "Thursday", "6" = "Friday", "7" = "Saturday")) + 
                    theme(axis.text.x = element_text(angle = 90)) + 
                    theme(legend.position= c(0.15,-0.24), legend.direction="horizontal") +
                    scale_fill_discrete(name = "Violent?") +
                      labs(title = "Number of reported robberies and thefts by weekday and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      ylab("Number of reports") + 
                    theme(axis.text.x = element_text(angle = 90)) + 
                    theme(legend.position= c(0.15,-0.235), legend.direction="horizontal") +
                    scale_fill_discrete(name = "Violent?") +
                    theme(axis.title.x = element_blank()) +
                    theme(strip.text = element_text(size = 6))
```

#### On public transportation, rhefts are most likely during peak commuting times. In cars, evenings are the most dangerous times for robberies.

On the Metro and Metrobus, times when thefts are most frequently reported match peak commuting times: Between 7-9am, 12-2pm and 6-8pm. Pesero crime reports have less data but seem to follow a similar pattern. For cars, on the other hand, reported robberies are consistently higher in the evenings, between 6-8pm.

```{r, warning = FALSE, message = FALSE, echo = FALSE}
 transport_crimes %>%  select(transp_mode, datetime, violence) %>% 
                      mutate(hour = hour(datetime), year = year(datetime)) %>% 
                     ggplot(aes(x = hour, fill = violence)) + geom_bar() +
                    facet_grid(year~transp_mode) +
                    theme_minimal() +
                    theme(strip.text.x = element_text(size = rel(0.6))) +
                    ylab("Number of crimes") + 
                    theme(axis.title.x = element_blank()) +
                    theme_minimal() + 
                      labs(title = "Number of reported robberies and thefts by mode, hour of day and year",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      ylab("Number of reports") + 
                      scale_x_continuous(breaks = c(0,5,10,15,20), labels= c(`0` = "00:00", `5` = "5:00", 
                                                                             `10` = "10:00", `15`= "15:00", `20` = "20:00")) + 
                    theme(axis.text.x = element_text(angle = 90)) +
                    theme(axis.title.x = element_blank()) + theme(axis.text=element_text(size=6)) + 
                    theme(legend.position= c(0.15,-0.11), legend.direction="horizontal") +
                    scale_fill_discrete(name = "Violent?")


```



#### Taking a closer look at crimes reported on specific modes shows some geographical variations

**In the metro system,  Hidalgo, Pantitlán, and Pino Suarez are the stations with the highest volume of reports of both robberies and thefts**

For thefts, Hidalgo, Pino Suárez, Pantitlán, Zócalo, and Chabacano. For robberies (that include violence) the top five stations where more crimes are reported are Hidalgo, Pantitlan, Pino Suárez, Tacuba and Zócalo. As a result, the neighborhoods where these stations are located (Centro, Guerrero, Pantiltlán) are the neighborhoods with the most reported crimes on metro stations.


```{r, warning = FALSE, message = FALSE, echo = FALSE, results='hide'}
# Load neighborhood ("colonias") shapefiles
colonias <- st_read("../data/coloniascdmx.shp", quiet = TRUE) %>% select(name = nombre, geometry) %>% st_transform(4485)
```


```{r, warning = FALSE, message = FALSE, echo = FALSE, results='hide'}
# make lat and lon into gemetrical object sf can work with
transport_crimes_geolocated <- transport_crimes %>% filter(!is.na(lat)) %>% st_as_sf(coords = c("lon", "lat"), crs=4326) %>% st_transform(4485)
```

```{r, warning = FALSE, message = FALSE, echo = FALSE, results='hide'}
metro_lines <- st_read("../data/lineas-de-metro/lineas-de-metro.shp", quiet = TRUE) %>% st_transform(4485) %>% 
                    mutate(line = fct_recode(name, "Line 1" = "Línea 1", 
                                               "Line 2" = "Línea 2",
                                               "Line 3" = "Línea 3",
                                               "Line 4" = "Línea 4",
                                               "Line 5" = "Línea 5",
                                               "Line 6" = "Línea 6",
                                               "Line 7" = "Línea 7",
                                               "Line 8" = "Línea 8",
                                               "Line 9" = "Línea 9",
                                               "Line A" = "Línea A",
                                               "Line B" = "Línea B",
                                               "Line 12" = "Línea 12")) %>% 
                                              select(-name)
                           
metro_stations <- st_read("../data/estaciones-metro/estaciones-metro.shp", quiet = TRUE) %>% st_transform(4485) %>% st_buffer(dist = units::as_units(250, "m"))

```


```{r, warning = FALSE, message = FALSE, echo = FALSE}

metro_crimes <- transport_crimes_geolocated %>% filter(transp_mode == "Metro")
cols <- c("Line 1" = "hotpink1", "Line 2" = "blue", "Line 3" = "olivedrab", 
          "Line 4" = "cadetblue1", "Line 5" = "yellow", "Line 6" = "red",
          "Line 7" = "orange", "Line 8" = "darkgreen", "Line 9" = "brown", 
          "Line A" = "purple", "Line B" = "grey69", "Line 12" = "darkgoldenrod")

violence.labs <- c(Yes = "Violent", No = "Non-violent")


ggplot(colonias) + 
  geom_sf(size = 0.1) + 
  geom_sf(data = metro_lines, aes(color = line), size = 1) + scale_colour_manual(values = cols) + 
  geom_sf(data=metro_crimes,  color = "black", size = 0.4, alpha = 0.5) +
    facet_wrap(~violence, labeller = labeller(violence = violence.labs)) +
    theme_void() + theme(legend.title = element_blank()) +
                      labs(title = "Locations of reported robberies and thefts on the Metro",
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México") +
                      theme(legend.position= "bottom", legend.direction="horizontal") 
```

***Car Robberies are reported in many areas of the city**

The neighborhoods with the most reported car robberies are Tacubaya, Anahuac, Doctores, San Pedro de los Pinos, Daniel Garza and Morelos. Most of the reported robberies are reported in neighborhoods in the Álvaro Obregon, Miguel Hidalgo and Iztapalapa municipalities.

```{r, warning = FALSE, message = FALSE, echo = FALSE}
# Function to count number of reports by neighborhood and prepare for map 
mode_crimes_counts <- function(mode){
  #filter out the reports of the chosen mode
mode_crimes <- transport_crimes_geolocated %>% filter(transp_mode == mode)
# spatial join with the neihborhoods data
colonias_crimes <- colonias %>% st_join(mode_crimes)
# count reports by neighborhood
colonias_counts <- colonias_crimes %>% 
  filter(!is.na(name)) %>% 
  st_drop_geometry() %>% 
  group_by(name) %>% 
  summarize(n = n()) %>% 
  full_join(colonias %>% 
              filter(!is.na(name)) %>%
              group_by(name) %>% 
              slice(1L) %>% 
              select(name, geometry)) %>%
  mutate(n = ifelse(is.na(n), 0, n)) %>%
  st_as_sf()
return(colonias_counts)
}
```


```{r, warning = FALSE, message = FALSE, echo = FALSE}
# Function to count top 10 neighborhoods for a given mode
top_10_neighborhoods <- function(mode){
  top_10 <- mode_crimes_counts(mode) %>% st_drop_geometry() %>% arrange(desc(n, by_group = TRUE)) %>% top_n(10)
  return(top_10)
}
```


```{r, warning = FALSE, message = FALSE, echo = FALSE}

# Function to map neighborhood reports for a given mode
mode_crimes_map <- function(mode){
  colonias_counts <- mode_crimes_counts(mode)
colonias_counts %>% ggplot(aes(fill = n)) + 
  geom_sf(size = 0.1) + 
  scale_fill_gradient(low = "white", high = "red", na.value = NA, trans = "log", breaks = c(1, 30, 100))  + 
  theme_void() + 
   theme(legend.title = element_blank()) +
  labs(title = paste0("Number of reported robberies and thefts by ", mode, " users"),
                      subtitle = "Mexico City 2016-2019",
                      caption = "Source: Carpetas de Investigacion PGJ, Ciudad de México")
}
```


```{r, warning = FALSE, message = FALSE, echo = FALSE}
mode_crimes_map("Car")
```

**Thefts and robberies reported on Peseros are laregly concentrated in neigborhoods en the east, with a large prportion in the Iztapalapa municipality**

The neighborhoods with the hightest number of reports include Ermita Zaragoza, Ejercito Constitucionalista, Santiago Atzacoalco, Martin Cabrera and Santa Marta Acatitla.

```{r, warning = FALSE, message = FALSE, echo = FALSE}
mode_crimes_map("Pesero")
```



#### The details of this analysis can be found in this Appendix on the data and this Appendix on visualization

For a detailed description of how the data was processed, a more thurough discussion of the limitations of this post and full `R` code for cleaning and visualizing the data used in this analysis check these additional posts.
